---
title: "Data Products - Quiz 1"
author: "Jose A. Ruiperez Valiente"
date: "7 de febrero de 2016"
output: html_document
---

# Question 1
Consider the following code for the cars data set

```{r}
library(manipulate)
myPlot <- function(s) {
    plot(cars$dist - mean(cars$dist), cars$speed - mean(cars$speed))
    abline(0, s)
}
```
This function plots distance versus speed, each de-meaned and an associated line of slope s.

Which of the following code will make a manipulate plot that creates a slider for the slope?

*Correct Answer:*

```{r}
manipulate(myPlot(s), s = slider(0, 2, step = 0.1))
```

# Question 2
Which of the following code uses the rCharts package to create a sortable and searchable data table for the airquality data set? Assume the rCharts package and the airquality data set have already been loaded into R.

*Correct Answer:*

```{r}
dTable(airquality, sPaginationType = "full_numbers")
```

# Question 3

A basic shiny data product requires:

*Correct Answer:*

**A ui.R and server.R file or a A server.R file and a directory called www containing the relevant html files.**

# Question 4

What is incorrect about the followig syntax in ui.R?

```{r}
library(shiny)
shinyUI(pageWithSidebar(  
  headerPanel("Data science FTW!"),  
  sidebarPanel(    
     h2('Big text')    
     h3('Sidebar')  
  ),  
  mainPanel(      
       h3('Main Panel text')  
  )
))
```

*Correct Answer*

**Missing a comma in the sidebar panel**

# Question 5

Consider the following code in ui.R

```{r}
shinyUI(pageWithSidebar(  
   headerPanel("Example plot"),  
   sidebarPanel(    
     sliderInput('mu', 'Guess at the mu',value = 70, min = 60, max = 80, step = 0.05,)  ), 
   mainPanel(    
     plotOutput('newHist')  
   )
))
```

And the following server.R code.

```{r}
library(UsingR)
data(galton)

shinyServer(  
    function(input, output) {    
       output$myHist <- renderPlot({      
          hist(galton$child, xlab='child height', col='lightblue',main='Histogram')      
          mu <- input$mu      
          lines(c(mu, mu), c(0, 200),col="red",lwd=5)      
          mse <- mean((galton$child - mu)^2)      
          text(63, 150, paste("mu = ", mu))      
          text(63, 140, paste("MSE = ", round(mse, 2)))      
          })      }
)
```

*Correct answer:*

**The server.R output name isn't the same as the plotOutput command used in ui.R.**